---
title: "2.Prediction"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
#IMPORTANT NOTICE:

##It is necessary to start predictions generating the classification models. DO it oppening the file **1.Training _Model.rmd** and follow instructions. If this is already done, follow instructions above.  


## 4. Predictions

The first step is to install knitr and libraries necessaries to do provenance predictions using the results from the generated models

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('SRC/FUNCTIONS/FirstConfiguration.R', echo=TRUE)
```

The actions above are needed to charge the results from the trained models, some functions needed to make predictions and to charge the databe with unknown provenance samples.

```{r}

load("OUTPUT/Trained_Model.RData")

source('SRC/FUNCTIONS/STACK_FUNCTION.R')
source('SRC/FUNCTIONS/WRITE_FUNCTION.R')

ruta<-"INPUTS/DATABASE 2 PREDICTION.xlsx"

PT <- read_excel(ruta, sheet = 1)
PT_txt<-PT[,2]
PT[,2]<-NULL

BASE<-PT
BASE$group<-NULL

head(read_excel(ruta, sheet = 1))

```

The head of the data from the document: **DATABASE 2 PREDICTION** can be seen on the results  above.

The next chunk code will make de predictions but the results will not be displayed below. 

```{r}
model_list<-list(GLM=BASE.lrm,RF=BASE.rf1,NNET=BASE.nnet,KNN=BASE.kn1,LDA=BASE.lda)

BASE_STACK<-stack_function(model_list,BASE,inf_group=FALSE)
predicciorf1<-predict(object=BASE_STACK.rf,newdata = BASE_STACK)
probstak<-predict(object=BASE_STACK.rf,newdata=BASE_STACK,type="prob")

BASE_STACK_PROB<-stack_function_prob(model_list,BASE,inf_group=FALSE)



BASE_STACK_PROB_TOTAL<-as.data.frame(c(BASE_STACK[,!is.element(names(BASE_STACK),"group")],BASE_STACK_PROB))

```

To produce the results, the following action will generate a .csv document with them.

```{r}


BASE_IMPRIMIR<-bbdd_print(PT_txt,BASE_STACK_PROB_TOTAL,predicciorf1,probstak)
imprimeix(nom="prediction_results.csv",df=BASE_IMPRIMIR)

```

The results are now available on your computer, you will find a .csv document on the folder: /Supervised_Provenance_Analysis/OUTPUT

Have a look at them and Good luck!

